---
title: "INFO 550 Homework 1 Solutions"
author: "Scott Gillespie"
date: "March 23, 2016"
output: html_document
---
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#url<-"http://steviep42.bitbucket.org/YOUTUBE.DIR/restaurants.db"
setwd("C:/Users/sgille3/Projects/Course Work/INFO - Software Engineering/Homework")
#download.file(url, "restaurants.db")
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(DBI)
library(RSQLite)
hwDB <- dbConnect(RSQLite::SQLite(), dbname="restaurants.db")
```

<h2>PART I:</h2> 

Solutions for SQLite questions #1 - 7 (copied from SQLite database)

The proceeding chunk performs the following tasks:

1.) Creates a character string containing SQL code for each question. \
2.) Pastes the character string for each question into a single list named 'sqlList' using a for loop.\
3.) Utilizes a function ("sqlFunction") to calculate the results of an input SQL query using the 'dbFetch' and 'dbSendQuery' functions from the DBI package.\
4.) Outputs the results of each query into a list.

```{r, warning=FALSE, message=FALSE}
sqlQuestion1<-("SELECT COUNT(*) FROM violations
               WHERE description LIKE ('%vermin%');") 


sqlQuestion2<-("SELECT COUNT(*) FROM violations
                WHERE ((risk_category = 'High Risk') AND description NOT LIKE ('%vermin%'));")


sqlQuestion3<-("SELECT COUNT(*) FROM inspections
                WHERE ((date BETWEEN 20140401 AND 20140501) AND (type LIKE ('%routine%')));")


sqlQuestion4<-("SELECT business_id, score, date, type FROM inspections
                WHERE score != 'NA' 
                ORDER BY score ASC
                LIMIT 10;")


sqlQuestion5<-("SELECT DISTINCT postal_code FROM businesses
                WHERE LENGTH(postal_code) = 5 AND postal_code != '00000';")


sqlQuestion6<-("SELECT name, postal_code, COUNT(name) AS mex_count FROM businesses
                WHERE name LIKE ('%mex%')
                GROUP BY postal_code
                ORDER BY mex_count DESC;")


sqlQuestion7<-("SELECT businesses.business_id, name, score FROM businesses, inspections
                WHERE ((businesses.business_id = inspections.business_id) AND (name LIKE ('%mex%')) AND (score != 'NA')) 
                ORDER BY score ASC
                LIMIT 10;")

sqlList<-list()
for (i in 1:7){
  sqlList[i] <- list(get(paste0('sqlQuestion',i)))
}

sqlFunction<-function(x){
  dbFetch(dbSendQuery(hwDB, x))
}

output<-list()
for (i in 1:7){
  output[[i]] <- sqlFunction(sqlList[[i]])
}
output
```
\
Creating dplyr object tables from the SQLite database

```{r, warning=FALSE, message=FALSE}
library(dplyr)
hwDBdplyr   <- src_sqlite(path="restaurants.db")
businesses  <- tbl(hwDBdplyr, "businesses")
inspections <- tbl(hwDBdplyr, "inspections")
violations  <- tbl(hwDBdplyr, "violations")
```
\
Running each query as before, except now utilizing dplyr package functionalities

```{r, warning=FALSE, message=FALSE}
dplyrQuestion8 <- violations %>% 
                  filter(description %like% ("%vermin%")) %>% 
                  summarise(Count=n())
dplyrQuestion8


dplyrQuestion9 <- violations %>% 
                  filter(description %not like% ("%vermin%"), (risk_category == "High Risk")) %>% 
                  summarise(Count=n())
dplyrQuestion9


dplyrQuestion10 <- inspections %>% 
                   filter(between(date, '20140401', '20140501'), (type %like% ('%routine%'))) %>% 
                   summarise(Count=n())
dplyrQuestion10


dplyrQuestion11 <- head(inspections %>% 
                        select(business_id, score, date, type) %>% 
                        filter(score != 'NA') %>% 
                        arrange((score)),10)
dplyrQuestion11


dplyrQuestion12 <- head(distinct(businesses %>% 
                        select(postal_code) %>% 
                        filter(length(postal_code) == 5, (postal_code != '00000'))),50)
dplyrQuestion12


dplyrQuestion13 <- head(businesses %>% 
                        select(name, postal_code) %>% 
                        filter(name %like% ('%MEX%')) %>% 
                        group_by(postal_code) %>% 
                        summarise(mex_count=count(name)) %>% 
                        arrange(desc(mex_count)), 50)
dplyrQuestion13


dplyrQuestion14 <- head(inner_join(businesses, inspections,"business_id") %>% 
                              select(business_id, name, score) %>% 
                              filter(name %like% ('%MEX%'), (score != 'NA')) %>% arrange(score),10)
dplyrQuestion14
```
\

<h2>PART II:</h2>

A function to quickly pull and select variables from JSON-formatted data, utilizing a web-based API provided by the NY Times. The data was copied from the NY Times website and stored locally as a file called "hw1.JSON". This file is provided on github and the direct weblink is commented-out within the function.

```{r, warning=FALSE, message=FALSE}
library(XML)
library(dplyr)
library(jsonlite)

nytimes_movies <- function(){

  jsonAPI <- fromJSON("C:/Users/sgille3/Projects/Course Work/INFO - Software Engineering/Homework/hw1.json")
  #http://api.nytimes.com/svc/movies/v2/reviews/all.json?api-key=4ce21e5e6b299a7e95d2a14a12345f5e:12:74694642
  retdf <- jsonAPI$results %>% 
            select(display_title, mpaa_rating, critics_pick, thousand_best, headline, opening_date) %>% 
            rename(names=display_title)
  
  return(retdf)
}
nyt <- nytimes_movies()
names(nyt)

nyt %>% 
  select(names, critics_pick, headline)
```
\

<h2>PART III:</h2>

<h3>Plot 1:</h3>

Average flight delays in minutes for US carriers. Arrival delays are summarized in red, and departure delays are summarized in blue. The research hypothesis concerned if more regional airlines (e.g. Hawaiian Airlines, Alaska Airlines, and Virgin America) performed better than major carriers (e.g. Frontier Airlines, Southwest Airlines, and United Airlines). Evidence from the plot seems to indicate a possible association between flight carrier type (regional vs. national) and average flight delays.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12,fig.height=9}
library(nycflights13)
library(ggplot2)
library(plotly)

airlines$carrier <- as.character(airlines$carrier)

##joinData
plot1Join <- inner_join(flights, airlines,"carrier") %>% 
  na.omit() %>%
  select(year, dep_delay, arr_delay, carrier, name) %>%
  group_by(carrier, name) %>%
  summarize(n = n(), avgDepDelay = mean(dep_delay), sdDepDelay = sd(dep_delay),
            n = n(), avgArrDelay = mean(arr_delay), sdArrDelay = sd(arr_delay)) %>%
  mutate(seDepDelay = sdDepDelay/sqrt(n), seArrDelay = sdArrDelay/sqrt(n))
  
##depDelay
depDelay <- plot1Join %>% 
  select(name, carrier, avgDepDelay, seDepDelay) %>%
  rename(avgDelay = avgDepDelay, seDelay = seDepDelay) %>%
  mutate(Type='Departure')
    
##arrDelay
arrDelay <- plot1Join %>% 
  select(name, carrier, avgArrDelay, seArrDelay) %>%
  rename(avgDelay = avgArrDelay, seDelay = seArrDelay) %>%
  mutate(Type='Arrival')

##fullData

fullData <- rbind(depDelay,arrDelay)

ggplot(fullData, aes(x=reorder(name,-avgDelay),y=avgDelay, fill=Type)) + 
          geom_bar(position=position_dodge(), 
             color="black", stat = "identity", size=0.3) +
          geom_errorbar(position=position_dodge(0.9), width=0.2, size=0.3, 
                  aes(ymin=avgDelay-seDelay,ymax=avgDelay+seDelay)) +
          coord_cartesian(ylim=c(-10,25)) +
          scale_y_continuous(breaks=c(-10,-5,0,5,10,15,20,25)) +
            theme_bw() + 
    geom_abline(intercept=0,slope=0, size=1.25) +
    labs(x="Flight Carrier", y="Average Delay (Minutes)") +
    scale_color_discrete(labels=c("Arrival", "Depature")) +
    theme(axis.title = element_text(color="#000000", face="bold", size=18),
          axis.text.x = element_text(face="bold", color="#000000", size=16, angle=58, hjust=1),
          axis.text.y = element_text(face="bold", color="#000000", size=16),
          legend.text = element_text(face="bold", color="#000000", size=24),
          legend.title= element_blank(), legend.position =c(0.89,0.90),
          axis.ticks = element_line(size = 1.5),
          panel.border = element_rect(linetype = "solid", colour = "black", size = 1.25))

```
\
\
\

<h3>Plot 2:</h3>

Pearson correlations between arrival delay, flight speed, and number of seats and engines. Lighter blue indicates stronger associations; whereas, darker blue indicates weaker association between the variables. From the plot, number of seats seems to correlate highly with speed of the jet; likewise, number of engines has moderate correlation with seats and speed. Interestingly, arrival delay fails to associate strongly with any of these characteristics.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8, fig.height=8}

library(reshape2)
plot2Join <- inner_join(flights, planes,"tailnum") %>%
  na.omit() %>% 
  mutate(Engines=engines, Seats=seats, Speed=speed, Delay=arr_delay) %>% 
  select(Delay, Speed, Seats, Engines)

plot2Corr<-as.matrix(cor(plot2Join))

lowerHalf<-function(x){
  x[upper.tri(x)] <- NA
  return(x)
}
plot2Corr<-lowerHalf(plot2Corr)
meltPlot2Corr<-melt(plot2Corr)

ggplot(data=meltPlot2Corr, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile(color = "white") +
  theme_minimal() +
  coord_fixed() +
  theme(axis.title = element_blank(),
  axis.text.x = element_text(face="bold", color="#000000", size=16),
  axis.text.y = element_text(face="bold", color="#000000", size=16),
  legend.text = element_text(face="bold", color="#000000", size=16),
  axis.ticks = element_line(size = 1.5),
  legend.title= element_blank())

```
